Blaming the Typeless: \\ Scalable, Human-Centric Python Localization

### 2018-08-24 paper/main.tex
rewrite related work

\begin{figure}
\begin{lstlisting}
year = int(time.strftime("%Y"))
age = input("Enter your age")
print("You will be twice as old in:")
print(year + age)
\end{lstlisting}
\caption{
\label{fig-motex}
A program with multiple fault localizations (lines 1 and 2).}
\end{figure}

@@

\section{Related Work}
\fixme{This must all be rewritten or dropped, as it is directly copied from
my previous papers. It serves as a placeholder to indicate length and to
remind us of some Software Engineering bits we might cite.}

There are several fault localization techniques that rely much more heavily
on dynamic information\footnote{Note: we do not directly compare to these
state-of-the-art fault localization techniques for several reasons.  First,
the benchmarks used in this paper do not always have available the test
suites needed by dynamic techniques and, conversely, those used in the
previous work do not have the publicly available bug reports required
by our technique.  Additionally, we return an answer only when confident in
our result set and also employ a top list of all available results, while
previous work is typically evaluated using a score metric.  We consider the
techniques complimentary.}~\cite
{harrold05,Renieris03,cleve05,wang09}.
In general,
these techniques leverage differences between passing and failing program
executions.  While effective, this type of approach requires program traces
for not only the specific fault at hand but also from a comprehensive regression
test suite for comparison.
A hybrid approach that considers both dynamic
traces and static natural language similarity was recently proposed by
Medini \textit{et al.}~\cite{Medini11}.
In contrast to these approaches, our technique relies purely on static
information that is readily available in most commercial systems with minimal
additional developer effort.

Prabhakararao and Ruthruff \emph{et al.} performed two human studies
to gauge the effectiveness of an interactive fault localization tool
developed for end users with little to no
experience~\cite{Prabhakararao03,ruthruff05}.  The goal of their
studies was to evaluate the use of feedback when locating faults and
to generally study the process of fault localization, especially by
users with no expert domain knowledge of the source.  By comparison,
our human study also examines the fault localization process but for
the purpose of evaluating software quality metrics.  We are less
interested in the specific process and more concerned with the
resulting accuracy and the human intuitions about the code in
question.  Additionally, our human study is of a much broader scope and
thus we hope it is more generalizable.
%Additionally, their studies included 10 and 20
%participants respectively using a spreadsheet-based visualization tool.
%Our study of 61 humans looking at actual source code potentially provides
%more generalizable results and targets specifically the area of software
%maintenance completed by trained programmers.

Ashok \textit{et al.} propose a similar natural language search technique
in which users can match an incoming report to previous reports,
programmers and source code~\cite{Ashok09}.  By comparison, our technique
is more lightweight and focuses only on searching the code and the
defect report.

Jones \textit{et al.} developed Tarantula, a technique that performs
fault localization based on the insight that statements executed often
during failed test cases likely account for potential fault
locations~\cite{harrold05}. Their approach is quite effective when
a rich, indicative test suite is available and can be run as part of
the fault localization process. It thus requires the fault-inducing
input but not any natural language defect report. By contrast,
our approach is lightweight, does not require an indicative test
suite or fault-inducing input, but does require a natural language
defect report. Both approaches will yield comparable performance, and
could even be used in tandem.

Cleve and Zeller localize faults by finding differences between
correct and failing program execution states, limiting the scope of
their search to only variables and values of interest to the fault in
question~\cite{cleve05}. Notably, they focus on those variable and
values that are relevant to the failure and to those program execution
points where transitions occur and those variables become causes of
failure. Their approach is in a strong sense finer-grained than ours:
while nothing prevents our technique from being applied at the level
of methods instead of files, their technique can give very precise
information such as ``the transition to failure happened when $x$
became 2.'' Our approach is lighter-weight and does not require
that the program be run, but it does require defect reports.

Renieris and Rice use a ``nearest neighbor'' technique in their
Whither tool to identify faults based on exposing differences in
faulty and non-faulty runs that take very similar executions
paths~\cite{Renieris03}. They assume a large number of correct runs
(e.g., normal test cases) and one failing run. Their approach uses a
distance criterion to select the correct run that is closest to the
failing run and produces a report of ``suspicious'' parts of the
program. By comparison, we chose to limit the programmatic information
used by our technique to only that which was reported by users: we
do not use test case runs but do need natural language.

Liblit \textit{et al.} use Cooperative Bug Isolation, a statistical
approach to isolate multiple bugs within a program given a deployed user
base. By analyzing large amounts of collected execution data from real
users, they can successfully differentiate between different causes of
faults in failing software~\cite{liblit05}. Their technique produces
a ranked list of very specific fault localizations (e.g., ``the fault
occurs when $i > arrayLen$ on line 57''). In general, their technique
can produce more precise results than ours, but it requires a set of
deployed users and works best on those bugs experienced by many users.
By contrast, we do not require that the program be runnable, much less
deployed, and use only natural language defect report text.

>>>>

\begin{figure}
\begin{lstlisting}
year = int(time.strftime("%Y"))
age = input("Enter your age")
print("You will be twice as old in:")
print(year + age)
\end{lstlisting}
\caption{
\label{fig-motex}
A simple program with multiple fault localizations (lines 1 and 2).}
\end{figure}

@@

\section{Related Work}

Broadly, the two most relevant areas of related work are 
software engineering approaches to fault localization (typically based on
dynamic test information) and programming languages approaches to fault
localization (typically based on static type information).

A significant body of work in fault localization follows from the Tarantula
project~\cite{tarantula}. Jones et al. proposed that statements executed
often in failing test runs but rarely in successful test runs were likely
to be implicated in the defect. Such spectrum-based approaches gather
coverage or other dynamic information and rank statements by a
mathematically-computed suspiciousness score. Subsequent algorithms
such as Pinpoint~\cite{chen02} or Ochiai~\cite{Abreu2006} have improved
upon Tarantula in various ways. For example, the Cooperative Bug Isolation
approach gathers richer dynamic information (i.e., invariants) and also
combines information from multiple end-user sources~\cite{liblit05}. 
Spectrum-based fault localization has been
the subject of significant empirical evaluation~\cite{Abreu2009}. 
Research on spectrum-based methods continues to this day, with
refinements, and even optimality results, to the associated
mathematical formulae~\cite{yoo2017}. While human studies have demonstrated
that users dislike large ranked lists~\cite{orso-parnin}, fault
localization has only increased in relevance with the rise of automated
program repair~\cite{genprog-tse-journal,semfix13}, where many techniques
depend critically on accurate fault localization~\cite{gpfaultloc}.

Our approach shares the overall goal of spectrum-based fault localization:
pinpointing suspicious statements to reducing debugging effort. However,
where spectrum-based traditionally focus on larger, industrial-scale 
programs, we target beginner-written software. Spectrum methods require
multiple test cases (ideally very many of them); we use just one program
execution. Spectrum methods focus heavily on dynamic features; we make
critical use of syntactic, contextual and type information as well. 
In addition, some human studies have focused specifically on accuracy and
expertise for fault localization~\cite{fry2010,Prabhakararao03,ruthruff05};
our decision to use particular features to support beginning programmers is
informed by such insights.

Zeller's popular Delta Debugging algorithm~\cite{zeller99}, interpreted
generally, efficiently finds a minimal ``interesting'' subset from among a
large set of elements. When the set of elements represents changes made to
a source code version control system and interesting is defined with
respect to failing a test suite, it can quickly locate program edits that
cause regressions. Alternatively, when considering correct and failing
execution states, such approaches can help focus on variable values that
cause failures~\cite{zeller05}. While our dynamic slicing information can
be viewed as a coarse approximation to the precise, fine-grained
localization such an approach can provide, a key difference is our use
of machine learning to agree with human judgments in cases where multiple
causes are equally logically valid.

In the programming languages community, a large body of work has focused on 
localizing faults and providing better error messages, typically through
the use of type information. \fixme{Sherrloc, Mycroft, Nate, etc.}
